---
author: "Bin Ma"
output: html_document
layout: post
title:  "Beeswarm plot with ggplot2"
subtitle: "ggbeeswarm package"
date: 2019-05-13  
background: '/img/posts/01.jpg'
---
**Beeswarm plots** are a way of plotting points that would ordinarily overlap so that they fall next to each other instead. In addition to reducing overplotting, it helps visualize the density of the data at each point (similar to a violin plot), while still showing each data point individually.

## Installation

```
install.packages('ggbeeswarm')
```

## Examples

### geom_quasirandom()

$geom_quasirandom$: Uses a van der Corput sequence or Tukey texturing (Tukey and Tukey "Strips displaying empirical distributions: I. textured dot strips") to space the dots to avoid overplotting.

```{r}
set.seed(12345)
library(ggplot2)
library(ggbeeswarm)
ggplot(iris, aes(Species, Sepal.Length)) + 
        geom_quasirandom()
```

Plotting with various width:

```{r}
ggplot(iris, aes(Species, Sepal.Length)) + 
        geom_quasirandom(varwidth = TRUE)
```

Plotting with fixed width:

```{r}
ggplot(iris, aes(Species, Sepal.Length)) + 
        geom_quasirandom(dodge.width=1)
```

### Alternative mthods

$geom_quasirandom$ can also use several other methods to distribute points. For example:

- Tukey

```{r}
ggplot(iris, aes(Species, Sepal.Length)) + geom_quasirandom(method = "tukey") + 
    ggtitle("Tukey texture")
```

- TukeyDense

```{r}
ggplot(iris, aes(Species, Sepal.Length)) + geom_quasirandom(method = "tukeyDense") + 
    ggtitle("Tukey + density")

```

- forwney

```{r}

ggplot(iris, aes(Species, Sepal.Length)) + geom_quasirandom(method = "frowney") + 
    ggtitle("Banded frowns")

```

- smiley

```{r}

ggplot(iris, aes(Species, Sepal.Length)) + geom_quasirandom(method = "smiley") + 
    ggtitle("Banded smiles")

```

- pseudorandom

```{r}

ggplot(iris, aes(Species, Sepal.Length)) + geom_quasirandom(method = "pseudorandom") + 
    ggtitle("Jittered density")
```

### geom_beeswarm

```{r}
ggplot(iris,aes(Species, Sepal.Length)) + 
        geom_beeswarm()
```


```{r}

ggplot(mpg,aes(class, hwy)) + 
        geom_beeswarm()

```

- With categorical y-axis

```{r}
ggplot(mpg,aes(hwy, class)) + 
        geom_beeswarm()
```

- ggplot doesn't pass any information about the actual device size of the points to the underlying layout code, so it's important to manually adjust the `cex` parameter for best results

```{r}
ggplot(mpg,aes(class, hwy)) + 
        geom_beeswarm(cex=5)
```

- pseudorandom

```{r}
ggplot(iris,aes(Species, Sepal.Length)) +       
        geom_beeswarm(cex=4,priority='density')
```

With automatic dodging:

```{r}
ggplot(iris, aes(Species, Sepal.Length)) + 
        geom_beeswarm(dodge.width=0.7, cex=4)
```

Copy from: Erik Clarke and Scott Sherrill-Mix